### 2021-12-15 report.Rmd
Update report.Rmd

---
title: "Report"
output: pdf_document
---

## Introduction

In this paper we replicate and extend an analysis published in JAMA Network Open by Benjamin P. Brown, MD, MS; Luciana E. Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  Brown et al. use a binary classification of state legislative climates on abortion as either "highly restrictive" or not, and they assess the association with abortion rates using propensity weighted linear difference in differences models, based on longitudinal data from 18 states.

We first reproduce the analysis of Brown et al. using data from the same 18 states.  The data set from the Brown et al. article was not made public, and due to limitations of publicly available data, we have narrowed the scope of our replication to analyzing a single year. Specifically we analyzed data from 2010, primarily because this was the year within the original paper's time range for which we had the most detailed information on state legislation restricting abortions.  We then expand the analysis by using data from all 50 states, and by using an integer variable for the extent of abortion restriction legislation.

The original study addresses the immediate, intended consequences of the restrictive abortion legislation.  In our paper, we extend the analysis to include associations of restrictive abortion legislation with fetal mortality, infant mortality, birth rate and fertility rate.  Restricted access to abortion has been linked to adverse outcomes for mothers and children, including pre-term birth and low birth weight (Redd et al., 2021), infant mortality rates (Edwards, 2019), and maternal mortality (Stevenson, 2021).  However, the effects of abortion legislation on these health outcomes are less well understood than the direct effect on number of abortions.  The key findings of a meta-analysis on the effects of targeted restrictions on abortion providers (TRAP) were that "Certain TRAP laws may be associated with a decrease in state-level abortion rates" and "Additional research is required to better understand the impact of these laws on women’s health outcomes." (Austin and Harper, 2017)  We attempt to address this needd in the existing literature by adapting the methodology of Brown et al. to quantify the health impacts of restrictive abortion legislation.  We also measure the association of these laws on birth rates, as restrictions on abortion have in some cases been promoted as an antedote to declining birth rates, recently in eastern Europe (Nikolova, 2020).

## Study Design

In our replication, we used propensity weighted linear models to assess the association of restrictive abortion legislation and abortion rates, using data from the same set of 18 states as Brown et al. (see Table 1).  We used propensity weighting to account for differences between those states and the other 32 statesbased on the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008. We fit linear models with and without distances to abortion providers and an interaction term between those distances and restrictive legislation.

## Participants, Variables and Data Sources

We used abortion count data compiled by Wm. Robert Johnston and available at www.johnstonsarchive.net/policy/abortion.  Abortion rates were calculated as number of abortions per female population. At least partial data was available for all 18 of the states included in the analysis by Brown et al., and for most of those states there was data for more than 80% of the counties (see Table 1).  The percentages of counties reporting were similar to the percentages reporting in Brown et al. (see their Table 1).

For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, TRAP laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors. Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3/4 types of laws.

We estimated distances to providers as the straight-line distances from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).

## Results

The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).  This suggests that the 18 states included in the analysis are not exceptionally different from those not included (in terms the demographic covariates listed above), and accordingly we expect the propensity weights to accurately adjust for the differences.

Tables 3 and 4 show the main results from our replication.  Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a  reduction in abortions of 0.814 per 1,000 women, and that effect was reduced to 0.649 when adjusting for distances to providers.  These effects were roughly twice as large as the effects calculated by Brown et al., nevertheless our estimates were within their reported 95% confidence intervals (see Brown et al. Table 3).

## Limitations

The original analysis and our replication are limited by the use of data from only 18 states.  We address this shortcoming in our extension below.  The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.

## Sources
Austin N, Harper S. "Assessing the impact of TRAP laws on abortion and women’s health in the USA: a systematic review" BMJ Sex Reprod Health, 2018; 44:128–134.

Edwards, E. "States pushing abortion bans have higher infant mortality rates". NBC News Website, 2019. Available at https://www.nbcnews.com/health/womens-health/states-pushing-abortion-bans-have-higher-infant-mortality-rates-n1008481. Accessed 12/14/21.

Nikolova, M. "When will conservatives learn that banning abortion does little to address demographic decline?" Emerging Europe website, 2020.  Available at https://emerging-europe.com/after-hours/when-will-conservatives-learn-that-banning-abortion-does-little-to-address-demographic-decline/. Accessed 12/14/21.

Redd, Sara K.; Stidham Hall, Kelli; Aswani, Monica S.; Sen, Bisakha; Wingate, Martha; Rice, Whitney S.  "Variation in Restrictive Abortion Policies and Adverse Birth Outcomes in the United States from 2005 to 2015" Women's Health Issues, 2021 1-11.

Stevenson, A. J. "The Pregnancy-Related Mortality Impact of a Total Abortion Ban in the United States: A Research Note on Increased
Deaths Due to Remaining Pregnant" Demography (2021) 58(6):2019–2028.