### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

---
title: "paper_replication"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load packages
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(geosphere)
library(stringr)
library(kableExtra)
library(table1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load in relevant data
providers <- read.csv("data/provider_list_geo.csv")
countycentroids <- read.csv("data/county_pop_centroids.csv")
countydem <- read.csv("data/county_demographics.csv")
abortions <- read.csv("data/cleanedabortiondata.csv")
restrictions <- read.csv("data/state_restrictions2010.csv")
countypres <- read.csv("data/countypres.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### clean county dem data ###########
countydem <- read.csv("data/county_demographics.csv")

for (i in 1:nrow(countydem)) {
  countydem$state[i] <- unlist(strsplit(countydem$NAME[i], ", "))[length(unlist(strsplit(countydem$NAME[i], ", ")))]
}

countydem$county <- gsub(countydem$NAME, pattern = ", [[:print:]]*$", replacement = "")
countydem$county <- gsub(countydem$county, pattern = " County", replacement = "")

countydem <- countydem %>% filter(countydem$state %in% state.name)

for (i in 1:nrow(countydem)) {
  countydem$state.abb[i] <- state.abb[which(state.name == countydem$state[i])]
}

countydem$county <- str_to_title(countydem$county)
countydem$county <- gsub(countydem$county, pattern = "Doña Ana", replacement = "Dona Ana")

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate abortion rates in 2010 ###########

#filter abortion rate data to only 2010
abortions2010 <- abortions %>% dplyr::select(state, county, X2010)

# length(unique(abortions2010$state)) #only 42 states in abortion data

#look at percent missing abortion counts by state
# aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

# clean abortion data to make counties and states match county dem data
abortions2010$county <- str_to_title(abortions2010$county)
abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
abortions2010$county <- gsub(abortions2010$county, pattern = "Ind. City", replacement = "City")
abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")

#fix NA values and make abortion counts numeric
abortions2010$X2010 <- na_if(abortions2010$X2010, "?")
abortions2010$X2010 <- na_if(abortions2010$X2010, "")
abortions2010$X2010 <- as.numeric(abortions2010$X2010)

#remove non-counties
abortions2010 <- abortions2010 %>% filter(!(county %in% c("Nonresidents", "Protected", "Other States", "Other Countries", "Non-Residents", "Other", "Nevada Protected", "Nevada Unspecified", "Nebraska Protected", "County", "Maryland")))

#look at counties in abortion data but not countydem data
# abortions2010$county[!(abortions2010$county %in% countydem$county)]
#"St. Louis County", "Bedford City", "Clifton Forge City"

#look at counties in countydem data but not abortion data
# countydem$county[!(countydem$county %in% abortions2010$county)] #many 
# unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"

#Add abortion counts to county data
for (i in 1:nrow(countydem)) {
  if (length(abortions2010$X2010[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]) == 0) {
    countydem$AbortionCount2010[i] <- NA
  }
  else {
    countydem$AbortionCount2010[i] <- abortions2010$X2010[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]
  }
}

#calculate abortion rates
countydem <- countydem %>% mutate(AbortionRate2010 = AbortionCount2010/Women)

#look at percent missing abortion rates by state
# aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate percent democratic votes by county in in 2008 ###########
countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)

countydem$county <- gsub(countydem$county, pattern = " Parish", replacement = "")
countydem$county <- gsub(countydem$county, pattern = " City", replacement = "")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " City", replacement = "")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Saint", replacement = "St.")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " County", replacement = "")

missingcounties <- countypresdem2008[!(countypresdem2008$county_name %in% countydem$county), ] #all of AK and a few others (DC, Dona Ana New Mexico, Kansas City Missouri, Maine Uocava Maine, Federal Precinct Rhode Island)

countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)

countydem <- left_join(countydem, countypresdem2008, by = c("state.abb", "county"))

#look at percent missing dem vote by state
# aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate distances from county centroids to closest provider location ###########

for (i in 1:nrow(countycentroids)) {
  lat <- countycentroids$LATITUDE[i]
  long <- countycentroids$LONGITUDE[i]
  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))*0.000621371
  countycentroids$distancetoprovider[i] <- min(dists)
}

countycentroids <- countycentroids %>% rename(county = COUNAME, state = STNAME)
countycentroids$county <- str_to_title(countycentroids$county)
countycentroids <- countycentroids %>% dplyr::select(county, state, distancetoprovider)
countycentroids <- countycentroids %>% filter(state != "Puerto Rico")
countycentroids$county <- gsub(countycentroids$county, pattern = " City", replacement = "")

missingcounties <- countycentroids[!(countycentroids$county %in% countydem$county), ] #Alaska and DC

countydem <- left_join(countydem, countycentroids, by = c("state", "county"))

#look at percent missing distance by state
# aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#make new variable for high vs. low restrictions
restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                              totalrestrict < 3 ~ "low"))
# restrictions$state[!(restrictions$state %in% countydem$state)] #none
countydem <- left_join(countydem, restrictions, by = "state")
countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#look at demographics

countydem <- countydem %>% mutate(inanalysis = case_when(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI") ~ 1, !(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI")) ~ 0))

countydemanalysis <- countydem %>% filter(inanalysis == 1)

table_1 <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
table_1 <- left_join(table_1, usedcounties, by = "state")
table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling(latex_options = 'hold_position')

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
##### Table 2 ####

countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$inanalysislabel <- factor(countydem$inanalysis, levels = c(1, 0), labels = c("In Analysis", "Not in Analysis"))

label(countydem$Women) = "Women"
label(countydem$White.pct) = "White"
label(countydem$Black.pct) = "Black"
label(countydem$Native.American.pct) = "Native American"
label(countydem$Asian.pct) = "Asian"
label(countydem$Hispanic.pct) = "Hispanic"
label(countydem$Median.Income) = "Median Income"
label(countydem$College.Grad.pct) = "College Graduates"
label(countydem$percentdem2008) = "Vote Democratic"

units(countydem$Women) = "No."
units(countydem$White.pct) = "%"
units(countydem$Black.pct) = "%"
units(countydem$Native.American.pct) = "%"
units(countydem$Asian.pct) = "%"
units(countydem$Hispanic.pct) = "%"
units(countydem$Median.Income) = "$"
units(countydem$College.Grad.pct) = "%"
units(countydem$percentdem2008) = "%"

# The following function is based on 
# "Using the table1 Package to Create HTML Tables of Descriptive Statistics"
# by Benjamin Rich.  Available at:
# https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-a-column-of-p-values
p_value_t_test <- function(x, ...){
    y <- unlist(x[1:2])
    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
    p <- t.test(y ~ g)$p.value  %>% round(3) 
    c("", sub("<", "<", format.pval(p, digits=3, eps=0.001)))
}

## Table 2 ###
table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
       data = countydem,
       overall = FALSE,
       extra.col = list(`P-value` = p_value_t_test),
       extra.col.pos = 3,
       footnote = 'P-values are derived by two sample t-tests.',
       caption = "County Data for States Included and Not Included in Analysis"
       )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=5, fig.align='center'}
#calculate propensity scores
propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
# summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score

# plot of distribution of propensity scores 
ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Figure 1: Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#run models
countydemanalysis <- df %>% filter(inanalysis == 1)

model_no_dist <- lm(AbortionRate2010 ~ restrictionlevel, data = countydemanalysis, weights = ps)
summ_no_dist <- summary(model_no_dist)

model_dist <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
summ_dist <- summary(model_dist)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Model Summaries (Note coefficients are per 1000 women)
table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3) %>%
  kable_styling(latex_options = 'hold_position')
table_3
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p <- summ_dist$coefficients[4,4] %>% round(3)
table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
                           model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_4

###### End of Replication ######
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Begin Extension #####
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("data/natalitydata.txt")
maternalmortality <- read.delim("data/maternalmortality.txt")
infantdeath <- read.delim("data/infantdeathrecords.txt")

infantdeath$County <- gsub(infantdeath$County, pattern = ", [[:print:]]*$", replacement = "")
infantdeath$County <- gsub(infantdeath$County, pattern = " County", replacement = "")
infantdeath$County <- str_to_title(infantdeath$County)
infantdeath$County <- na_if(infantdeath$County, "")
infantdeath <- infantdeath %>% filter(County != "Unidentified Counties")
infantdeath$County <- gsub(infantdeath$County, pattern = " Parish", replacement = "")
infantdeath$County <- gsub(infantdeath$County, pattern = "Borough", replacement = "Municipality")
infantdeath$County <- gsub(infantdeath$County, pattern = " City", replacement = "")
# infantdeath$County[!(infantdeath$County %in% countydem$county)]
infantdeath <- infantdeath %>% rename(state = State, county = County, infantdeathrate = Death.Rate)
infantdeath <- infantdeath %>% dplyr::select(state, county, infantdeathrate)

maternalmortality$County <- gsub(maternalmortality$County, pattern = ", [[:print:]]*$", replacement = "")
maternalmortality$County <- gsub(maternalmortality$County, pattern = " County", replacement = "")
maternalmortality$County <- str_to_title(maternalmortality$County)
maternalmortality$County <- na_if(maternalmortality$County, "")
maternalmortality <- maternalmortality %>% filter(County != "Unidentified Counties")
maternalmortality$County <- gsub(maternalmortality$County, pattern = " Parish", replacement = "")
maternalmortality$County <- gsub(maternalmortality$County, pattern = "Borough", replacement = "Municipality")
maternalmortality$County <- gsub(maternalmortality$County, pattern = " City", replacement = "")
maternalmortality$County <- gsub(maternalmortality$County, pattern = "La Porte", replacement = "Laporte")
maternalmortality$County <- gsub(maternalmortality$County, pattern = "De Kalb", replacement = "Dekalb")
maternalmortality$County <- gsub(maternalmortality$County, pattern = "Mc Kean", replacement = "Mckean")
maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "Municipality", replacement = "Borough")
maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "Borough", replacement = "And Borough")
maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "/", replacement = " ")
maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = " Census Area", replacement = "")
maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = " And", replacement = "")
# maternalmortality$County[!(maternalmortality$County %in% countydem$county)]

maternalmortality <- maternalmortality %>% rename(state = State, county = County, maternaldeaths = Deaths, maternaldeathrate = Crude.Rate)
maternalmortality <- maternalmortality %>% dplyr::select(state, county, maternaldeathrate)

natality$County <- gsub(natality$County, pattern = ", [[:print:]]*$", replacement = "")
natality$County <- gsub(natality$County, pattern = " County", replacement = "")
natality$County <- str_to_title(natality$County)
natality$County <- na_if(natality$County, "")
natality <- natality %>% filter(County != "Unidentified Counties")
natality$County <- gsub(natality$County, pattern = " Parish", replacement = "")
natality$County <- gsub(natality$County, pattern = "Borough", replacement = "Municipality")
natality$County <- gsub(natality$County, pattern = " City", replacement = "")
natality$County <- gsub(natality$County, pattern = "La Porte", replacement = "Laporte")
# natality$County[!(natality$County %in% countydem$county)]
natality <- natality %>% rename(state = State, county = County)
natality <- natality %>% dplyr::select(state, county, Birth.Rate, Fertility.Rate)

countydem <- left_join(countydem, infantdeath, by = c("state", "county"))
countydem <- left_join(countydem, maternalmortality, by = c("state", "county"))
countydem <- left_join(countydem, natality, by = c("state", "county"))

# sum(is.na(countydem$infantdeathrate)/nrow(countydem))
countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Missing")

# sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
# sum(is.na(countydem$Birth.Rate)/nrow(countydem))

```
## Expansion to Data from All States
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### recode restriction level and include all counties with abortion data in analysis ####
countyanalysis <- function(varname) {
  countydem <- countydem %>% mutate(hasdata = case_when(is.na(countydem[[varname]]) ~ 0, 
                                                              !(is.na(countydem[[varname]])) ~ 1))
  countydemanalysis <- countydem %>% filter(hasdata == 1)
  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
  usedcounties <- aggregate(countydem[[varname]] ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
  summarytable <- left_join(summarytable, usedcounties, by = "state")
  summarytable <- summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% 
    kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), 
        caption = "Number of Counties with Data Included in Analysis by State") %>% 
    kable_styling(latex_options = 'hold_position')

  countydem$hasdatalabel <- factor(countydem$hasdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))

  label(countydem$Women) = "Women"
  label(countydem$White.pct) = "White"
  label(countydem$Black.pct) = "Black"
  label(countydem$Native.American.pct) = "Native American"
  label(countydem$Asian.pct) = "Asian"
  label(countydem$Hispanic.pct) = "Hispanic"
  label(countydem$Median.Income) = "Median Income"
  label(countydem$College.Grad.pct) = "College Graduates"
  label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"

  units(countydem$Women) = "No."
  units(countydem$White.pct) = "%"
  units(countydem$Black.pct) = "%"
  units(countydem$Native.American.pct) = "%"
  units(countydem$Asian.pct) = "%"
  units(countydem$Hispanic.pct) = "%"
  units(countydem$Median.Income) = "$"
  units(countydem$College.Grad.pct) = "%"
  units(countydem$percentdem2008) = "%"

  # p_value_t_test <- function(x, ...){
  #   y <- unlist(x[1:2])
  #   g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
  #   p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
  #   p_value <- ifelse(p_value == '0', '< 0.001', p_value)
  # }
  # 
  # in_analysis_comp_table <- table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
  #      data = countydem,
  #      extra.col = list(`P-value` = p_value_t_test),
  #      footnote = 'P-values are derived by two sample t-tests.'
  #      )

  #calculate propensity scores
  #include education this time
  propensitywts <- glm(hasdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())

  p.score <- predict(propensitywts, countydem, type = "response")
  df <- countydem
  df$ps <- p.score

  #plot of distribution of propensity scores 
  # p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
  # xlab('Probability of Being in the Analysis') +
  # ggtitle('Propensity Score Distribution by Group') +
  # scale_fill_discrete('') +
  # theme(legend.position = 'bottom', legend.direction = 'vertical') +
  # ylab("Density")
  # print(p1)

  #run model
  countydemanalysis <- df %>% filter(hasdata == 1)
  model_no_dist <- lm(countydemanalysis[[varname]] ~ totalrestrict, data = countydemanalysis, weights = ps)

  model_dist <- lm(countydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
  return(list(summary(model_no_dist), summary(model_dist), summarytable))
}

returns <- countyanalysis("AbortionRate2010")
all_state_model_no_dist <- returns[[1]]
all_state_model_dist <- returns[[2]]

# Model Summaries (Note coefficients are per 1000 women)
table_5 <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = all_state_model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3) %>%
  kable_styling(latex_options = 'hold_position')
table_5

p <- all_state_model_dist$coefficients[4,4] %>% round(3)
table_6 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(all_state_model_dist$coefficients[2]*1000, 
                           all_state_model_dist$coefficients[3]*1000,
                           all_state_model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_6

all_state_table <- returns[[3]]
# all_state_table # move to appendix?
```

## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
```{r, echo=FALSE, message=FALSE, warning=FALSE}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
# hist(countydem$maternaldeathrate)

countydem$infantdeathrate <- na_if(countydem$infantdeathrate, "Missing County")
countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "[()]", replacement = "")
countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "Unreliable", replacement = "")
countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
# hist(countydem$infantdeathrate)

countydem$Birth.Rate <- na_if(countydem$Birth.Rate, "Missing County")
countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
# hist(countydem$Birth.Rate)

countydem$Fertility.Rate <- na_if(countydem$Fertility.Rate, "Missing County")
countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
# hist(countydem$Fertility.Rate)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Large County Function

largecountyanalysis <- function(varname) {
  largecountydem <- countydem %>% filter(Total >= 250000)
  largecountydem <- largecountydem %>% mutate(hasdata = case_when(is.na(largecountydem[[varname]]) ~ 0, 
                                                              !(is.na(largecountydem[[varname]])) ~ 1))
  largecountydemanalysis <- largecountydem %>% filter(hasdata == 1)
  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
  usedcounties <- aggregate(largecountydem[[varname]] ~ state, data = largecountydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
  summarytable <- left_join(summarytable, usedcounties, by = "state")
  # print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling())

  largecountydem$hasdatalabel <- factor(largecountydem$hasdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))

  label(largecountydem$Women) = "Women"
  label(largecountydem$White.pct) = "White"
  label(largecountydem$Black.pct) = "Black"
  label(largecountydem$Native.American.pct) = "Native American"
  label(largecountydem$Asian.pct) = "Asian"
  label(largecountydem$Hispanic.pct) = "Hispanic"
  label(largecountydem$Median.Income) = "Median Income"
  label(largecountydem$College.Grad.pct) = "College Graduates"
  label(largecountydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"

  units(largecountydem$Women) = "No."
  units(largecountydem$White.pct) = "%"
  units(largecountydem$Black.pct) = "%"
  units(largecountydem$Native.American.pct) = "%"
  units(largecountydem$Asian.pct) = "%"
  units(largecountydem$Hispanic.pct) = "%"
  units(largecountydem$Median.Income) = "$"
  units(largecountydem$College.Grad.pct) = "%"
  units(largecountydem$percentdem2008) = "%"

  # print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
  #      data = largecountydem,
  #     ))

  #calculate propensity scores
  #include education this time
  propensitywts <- glm(hasdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=largecountydem, family=binomial())

  p.score <- predict(propensitywts, largecountydem, type = "response")
  df <- largecountydem
  df$ps <- p.score

  #plot of distribution of propensity scores 
  p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
  # print(p1)

  #run model
  largecountydemanalysis <- df %>% filter(hasdata == 1)

  model_no_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict, data = largecountydemanalysis, weights = ps)

  model_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = largecountydemanalysis, weights = ps)

  return(list(summary(model_no_dist), summary(model_dist)))
}

returns <- largecountyanalysis("infantdeathrate")
large_county_no_dist <- returns[[1]]
large_county_dist <- returns[[2]]

# Model Summaries
table_7 <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = large_county_no_dist$coefficients[2],
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_7

p_2 <- large_county_dist$coefficients[3,4] %>% round(3)
p_3 <- large_county_dist$coefficients[4,4] %>% round(3)

table_8 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(large_county_dist$coefficients[2], 
                           large_county_dist$coefficients[3],
                           large_county_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_8

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# largecountyanalysis("Birth.Rate")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# largecountyanalysis("Fertility.Rate")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#add in state level maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####

statelevelnatality <- read.delim("data/statelevelnatality.txt")
statelevelinfantdeaths <- read.delim("data/statelevelinfantdeaths.txt")
statelevelfetaldeaths <- read.delim("data/statelevelfetaldeaths.txt")
statelevelmaternalmortality <- read.delim("data/statelevelmaternalmortality.txt")

statelevelnatality <- statelevelnatality %>% select(State, Birth.Rate, Fertility.Rate)
statelevelinfantdeaths <- statelevelinfantdeaths %>% select(State, Death.Rate)
statelevelfetaldeaths <- statelevelfetaldeaths %>% select(Standard.Residence.States, Percent.of.Total.Deaths) %>% rename(State = Standard.Residence.States, Pct.Fetal.Deaths = Percent.of.Total.Deaths)
statelevelmaternalmortality <- statelevelmaternalmortality %>% select(State, X..of.Total.Deaths) %>% rename(Pct.Maternal.Deaths = X..of.Total.Deaths)

statedem <- countydem %>% group_by(state) %>% summarize(AvgWomen = mean(Women, na.rm = TRUE), AvgWhite.pct = mean(White.pct, na.rm = TRUE), AvgBlack.pct = mean(Black.pct, na.rm = TRUE), AvgNative.American.pct = mean(Native.American.pct, na.rm = TRUE), AvgAsian.pct = mean(Asian.pct, na.rm = TRUE), AvgHispanic.pct = mean(Hispanic.pct, na.rm = TRUE), AvgCollege.Grad.pct = mean(College.Grad.pct, na.rm = TRUE), AvgMedian.Income = mean(Median.Income, na.rm = TRUE), AvgDem2008.pct = mean(percentdem2008, na.rm = TRUE), AvgDistancetoProvider = mean(distancetoprovider, na.rm = TRUE), TotalRestrictions = mean(totalrestrict, na.rm = TRUE), AvgAbortionRate2010 = mean(AbortionRate2010, na.rm = TRUE)) %>% rename(State = state)

statedem <- left_join(statedem, statelevelnatality, by = c("State"))
statedem <- left_join(statedem, statelevelinfantdeaths , by = c("State"))
statedem <- left_join(statedem, statelevelfetaldeaths, by = c("State"))
statedem <- left_join(statedem, statelevelmaternalmortality, by = c("State"))

for (i in 1:ncol(statedem)) {
  statedem[[i]][is.nan(statedem[[i]])] <- NA
}

statedem$Pct.Maternal.Deaths <- na_if(statedem$Pct.Maternal.Deaths, "Suppressed")
statedem$Pct.Maternal.Deaths <- gsub(statedem$Pct.Maternal.Deaths, pattern = "%", replacement = "")
statedem$Pct.Maternal.Deaths <- as.numeric(statedem$Pct.Maternal.Deaths)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
###make function for extended analysis###
stateanalysis <- function(varname) {
  statedem <- statedem %>% mutate(hasdata = case_when(is.na(statedem[[varname]]) ~ 0, 
                                                              !(is.na(statedem[[varname]])) ~ 1))
  statedemanalysis <- statedem %>% filter(hasdata == 1)
  statedem$hasdatalabel <- factor(statedem$hasdata, levels = c(1, 0), labels = c("States With Data", "States Without Data"))

  label(statedem$AvgWomen) = "Average Women"
  label(statedem$AvgWhite.pct) = "Average White"
  label(statedem$AvgBlack.pct) = "Average Black"
  label(statedem$AvgNative.American.pct) = "Average Native American"
  label(statedem$AvgAsian.pct) = "Average Asian"
  label(statedem$AvgHispanic.pct) = "Average Hispanic"
  label(statedem$AvgMedian.Income) = "Average Median Income"
  label(statedem$AvgCollege.Grad.pct) = "Average College Graduates"
  label(statedem$AvgDem2008.pct) = "Democratic Vote"

  units(statedem$AvgWomen) = "No."
  units(statedem$AvgWhite.pct) = "%"
  units(statedem$AvgBlack.pct) = "%"
  units(statedem$AvgNative.American.pct) = "%"
  units(statedem$AvgAsian.pct) = "%"
  units(statedem$AvgHispanic.pct) = "%"
  units(statedem$AvgMedian.Income) = "$"
  units(statedem$AvgCollege.Grad.pct) = "%"
  units(statedem$AvgDem2008.pct) = "%"

  # print(table1(~ AvgWomen + AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct | hasdatalabel, 
  #      data = statedem,
  #      extra.col = list(`P-value` = p_value_t_test),
  #      footnote = 'P-values are derived by two sample t-tests.'
  #      ))

  #calculate propensity scores
  #include education this time
  propensitywts <- glm(hasdata ~ AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct, data=statedem, family=binomial())

  p.score <- predict(propensitywts, statedem, type = "response")
  df <- statedem
  df$ps <- p.score

  #plot of distribution of propensity scores

  # plot1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
  #   xlab('Probability of Being in the Analysis') +
  #   ggtitle('Propensity Score Distribution by Group') +
  #   scale_fill_discrete('') +
  #   theme(legend.position = 'bottom', legend.direction = 'vertical') +
  #   ylab("Density")
  # print(plot1)

  #run model
  statedemanalysis <- df %>% filter(hasdata == 1)

  model_no_dist <- lm(statedemanalysis[[varname]] ~ TotalRestrictions, weights = ps, data = statedemanalysis)

  model_dist <- lm(statedemanalysis[[varname]] ~ TotalRestrictions + AvgDistancetoProvider + TotalRestrictions*AvgDistancetoProvider, data = statedemanalysis, weights = ps)

  return(list(summary(model_no_dist), summary(model_dist)))
}

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# returns <- stateanalysis("AvgAbortionRate2010")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### State Birth Rate
returns <- stateanalysis("Birth.Rate")
model_no_dist <- returns[[1]]
model_dist <- returns[[2]]

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_9 <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_9

p_1 <- model_dist$coefficients[2,4] %>% round(3)
p_2 <- model_dist$coefficients[3,4] %>% round(3)
p_3 <- model_dist$coefficients[4,4] %>% round(3)

table_10 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_10
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### State Fertility Rate
returns <- stateanalysis("Fertility.Rate")
model_no_dist <- returns[[1]]
model_dist <- returns[[2]]

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

p_1 <- model_dist$coefficients[2,4] %>% round(3)
p_2 <- model_dist$coefficients[3,4] %>% round(3)
p_3 <- model_dist$coefficients[4,4] %>% round(3)

table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### State  Infant Death Rate
returns <- stateanalysis("Death.Rate")
model_no_dist <- returns[[1]]
model_dist <- returns[[2]]

# p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = '< 0.001') %>% 
  kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

# p_1 <- model_dist$coefficients[2,4] %>% round(3)
p_2 <- model_dist$coefficients[3,4] %>% round(3)
p_3 <- model_dist$coefficients[4,4] %>% round(3)

table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

### State Fetal Death Percent
returns <- stateanalysis("Pct.Fetal.Deaths")
model_no_dist <- returns[[1]]
model_dist <- returns[[2]]

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

p_1 <- model_dist$coefficients[2,4] %>% round(3)
p_2 <- model_dist$coefficients[3,4] %>% round(3)
p_3 <- model_dist$coefficients[4,4] %>% round(3)

table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### State Maternal Deaths
returns <- stateanalysis("Pct.Maternal.Deaths")
model_no_dist <- returns[[1]]
model_dist <- returns[[2]]

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

p_1 <- model_dist$coefficients[2,4] %>% round(3)
p_2 <- model_dist$coefficients[3,4] %>% round(3)
p_3 <- model_dist$coefficients[4,4] %>% round(3)

table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

\newpage

## Sources

Data on the county population centroids came from the United States Census Bureau.
Available at https://www.census.gov/geographies/reference-files/time-series/geo/centers-population.html Accessed 12/11/21.

Data on the county-level voting totals came from the Harvard Dataverse.
Available at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ. Accessed 12/13/21.

## Code Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```